Re-Allocation of Resources during Releases \titlenote{Copyright note}

### 2013-12-04 paper1/paper1.tex
finished writing all sections.

\title{Re-Allocation of Resources during Releases \titlenote{Copyright note}}
\numberofauthors{2}
\author{
% 1st. author
\alignauthor
Md Tajmilur Rahman\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \affaddr{438-932-2288, +1}\\
       \email{mdt\_rahm@encs.concordia.ca}
% 2nd. author
\alignauthor
Peter C. Rigby\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \affaddr{514-848-2424, +1}\\
       \email{peter.rigby@concordia.ca}
}
\date{20 Oct 2013}
\maketitle
\begin{abstract}
This section will be written at the end.
\end{abstract}
\category{K.6.3}{Software Management}{Software Development, Software Resource Management, Resource Reallocation}
\terms{Experiment, Human Factors, Resource Management, Reallocation}
\keywords{Resource Reallocation, Code Ownership, Defect Density, Software Releases} % NOT required for Proceedings
\section{Introduction}
Software projects are notorious for going over budget and schedule. Rush periods are often get seen before a major release that turn the developers into dinosaurs as Frederick Brooks likens in his benchmark study "The Mythical Man Month" \cite{brooks_mythical}. This "Rush To Release (RTR)" can be prompted either by external forces such as decisions by management to include new features in the release or to release earlier to beat a competitor. Alternatively, the rush may simply be due to inappropriate or unrealistic scheduling. Whatever the reason is it is an obvious. Regardless of the causes, the rush to release stresses developers and often requires developers to work on unusual, high priority or critical areas of the system. In this paper we study how RTR may effect project organization and introduce technical debt. We want to propose a method to observe, analyze and summarize the results found around different parts of a release cycle. We intend to infer behavior of the developers related to the process around release time. The key research questions that we expect to answer with our methodology are described in section 1.1.

During the merge window, developers are free to provide any change that is ready for release. During the release candinate period, we would expect to see fixs on the files that were changed during the previous merge window. We measure this by comparing the Jaccard distance of merge window with the subsequent rc period, compared with the merge window and subsequent merge window.

@@

We have organized this paper as follows. In Section 2, we describe some background and motivations followed by the summaries of related works in Section 3. Section 4 will describe methodology of our work and explanation of the data we are working with for this research. In section 5 we will dicuss the results of the findings. Finally section 6 will conclude the paper.

@@

\renewcommand{\labelenumi}{RQ\theenumi:}
\begin{enumerate}
\item What is the release process used by the project? \newline
We will be giving a qualitative description of Linux Kernel development in brief to give anser to this question.
\item How many developers are allocated troughout different segments of a release cycle? What significant difference can be observed between the segments of a release cycle? \newline
Answer to this question will give us a big picture of Linux Kernel development community as well as the contrubution of developers in different parts within a stable release.
\item Do developers work on different areas of the system around the time of release? \newline
We want to see if developers are looking at different types of files in the code-base around the time of release which is being considered as the release period or rc period. We are going to introduce an algorithm to traverse through the git DAG to find out commits associated to a particular release. This can be useful for any other data collected from distributed version control system like git. We are expecting this would be the RTR in a release cycle which may introduce technical debts.
\item Do developers work in others' code in a high proportion during the rush period?\newline
Developers work on various files during release period. We will try to find out what is the percentage of dealing with files that developers own or they are most familier to. This will give us a sense of the nature of work during release time.
\item How long a commit takes to get pushed into the main kernel? Does it differ in different periods of release?\newline
In distributed development environment developers work in different branches. They commit their codes to the branches they are associated to. While core development is going on developers  do not directly push their changes to the main branch. We would like to see how much time it takes for a commit to the local branch to get pushed to the main branch. We also should understand how long it takes for regular development commits and how long it takes for the commits made around the release period.
\item Are there certain areas of the system that receive increased attention (i.e.\ do developers focus on a smaller set of files around releases)? \newline
To answer this question we will look into the files getting churned extremely high in number. We will try to figure out the difference between those highly churned or highly focused sets of files in different releases to understand if they require contantly high focus or they are only for one particular release.
\end{enumerate}

@@

\section{Related Works}
Hindle worked on release pattern discovery via partitioning \cite{hindle_release_pattern} to propose an approach to characterize a project's behavior around the time of major and minor releases while we are trying to study the behavior of the development resources around the time of release. In this research they proposed a method of observing, analyzing and summarizing the results of metrics of revisions found near releases. They have characterized a project's behavior around the time of major and minor releases. This is done by partitioning the observed activities like the artiffect check-ins around the dates of major and minor releases, then look for reasonable patterns. Hindle divided the revisions in each release in 4 different classes, Source Code, Testing, Building, and Documentations. On the other hand Cook did an interesting job, he inserted sensors and monitors into the development process but Hindle analyzed the data to understand what happened in the past \cite{cook_automating}. Basically Hindle worked in a reverse way than Cook did. Another research work we would like to mention was done by  Damian where they have worked on the role of domain knowledge and cross functional communication among the OOS development teams \cite{damian_domain}. Posnett did some dual ecological measures of focus in software development \cite{posnett_ecological}. Posnett's measure was for the more general view that unifies developer focus and artifact ownership. He analyzed i) developer artifact contribution to network to a predator-prey food web ii) drew upon ideas from ehology to produce a novel and iii) conceptually unified view of measuring focus and ownership. Another study was done by F. Rahman about the authorship of the code-bases in OSS development \cite{rahman_ownership}.

\section{Methodology and Data}
This section presents our methodology for discovering information which can give us the idea to get the answers to our research questions. We have collected the development history data of Linux kernel. It is a data source containing the the historical data of Linux kernel development since 2005. We are going to present the steps involved in this process and then we will follow up with an application of our methodology in a case study. In order to address our research questions, we obtain key measures of project evolution from the archival data collected from the distributed version contron (DVC) system of Git where all information of the OSS project ``Linux Kernel Development'' is recorded in electronic form. Many other OSS projects archive similar data, so the techniques used here can be replicated on any such project. We used the data elements extracted from the archival source to construct a number of measures on the commit log records to understand the behavior of the developers.

@@

\subsubsection{Explanation}
A Linux kernel development release version having maximum length of information does look like \textbf{linuxvA.B.C-rcP} where A, B, C, P are numeric. If a release versioning looks like ``linuxv2.6.13'' it tells us that this particular release is the 13th minor release  in the seriese of the 6th major release of kernel version 2. ``linuxv2.6.13\textbf{-rc1}'' says that after the 12th stable minor release had published, development for the next release has been started and rc1 is the first candidate on the way to the next release. It also ensures that next release is going to be another minor release which is ``linuxv2.6.13''. We have last couple of release candidates for ``linuxv2.6.12'' and first couple of release candidates for ``linuxv3.12'' which is not useful for our calculation. So we are having 340 releases including major, minor, micro and rc; 39 stable releases from ``linuxv2.6.11'' to ``linuxv3.11'' having complete data with us. We have total 400441 distinct commits and for 372943 of them have change information for 77082 different files in our hand where we have 14599 developers' working history among 14621 distinct developers. Surprisingly 14569 individual files didn't have any change (addition or deletion) but was being committed at least once. So we found them in our released wise datasets.

@@

\subsection{Finding Commits for Releases}
We have to sort out and accumulate all the commits together those are belonging to any of the releases. We have all 400441 commits in our hand but though we can't plot them into a single time domain to assign different set of commits to different release periods because development goes in parallel on different branches in git, so commits also go in parallel. If there are multiple branches, surely there are multiple merges and the commits prior to a crisscross merge are all ancestors of the heads of all subsequent branches \cite{bird_git}. Figure 1 shows how a git DAG looks like having so many crisscross merges in a busy branching snap-shot of the git DAG of Linux Kernel development.
\begin{figure}
\begin{center}
\includegraphics[height=2.5in,width=2.5in]{gitdag.png}
\caption{\small \sl Branching in git DAG for linux kernel development.}
\end{center}
\end{figure}
So it is possible to make a commit months ago from today while several releases may have already been made in the mean time but that commit can belong to any future release. Commits that are being made to a branch may get merged with other branch or the head branch during the release period. This is why we had to track all the way down the trees for each of the 340 releases in the git DAG to find out which commit contributes to which release. We could do this using Git DAG Traversing (GDT) algorithm as shown in Algorithm 1. After applying GDT we collected 381152 out of 400441 commits those are well organized in the git DAG of Linux Kernel development. As we mentioned earlier that we have last couple of release candidates for ``linuxv2.6.12'' and first couple of release candidates for ``linuxv3.12''. We are choping them out because we don't have complete commit records for version 2.6.12-rc releases as well as 3.12-rc releases. Out of 400441 commits 19289 are belonging to those releases. Therefore, finally we have 381152 commits for 331 releases in our database to progress our work on.

\begin{algorithm}
\caption{GDT: to find Commits within Releases}
\begin{algorithmic}[1]
\REQUIRE
\STATE commit id for all releases
\ENSURE
\FOR{all commits}
\STATE start from new commit, release R
\STATE store commit in ``git\_commit\_release''
\IF{has parent commit}
	\FOR{each parent commit}
		\IF{if parent commit has a release tag}
			\STATE repeat step 1
		\ELSIF{parent commit already stored}
			\STATE compare release versions
			\IF{release version not smaller than R}
				\STATE update existing\_release $\gets$ R
			\ENDIF
		\ELSE
			\STATE store parent commit in ``git\_commit\_release''
			\STATE commit id $\gets$ parent commit id
			\STATE repeat step 3
		\ENDIF
	\ENDFOR
\ELSIF{no parent commit}
	\STATE repeat step 1
\ENDIF
\ENDFOR
\end{algorithmic}
\end{algorithm}

\subsection{Time-Span of Releases}
Another information that we require is what are the durations of the releases of Linux kernel development. To achieve that we joined the table ``git\_refs\_tags'' where all the releases are stored including the splitted release numbers and the table ``git\_commit'' to give us all the dates of commits for each and every releases, so that we can easily find out the duration between two consecutive releases (including release candidates RC). This information will significantly help us because we need to see what is the development period and what are the release periods within a release cycle. We also need to understand how developers are working, what is the impact of their changes made in the code-base during a particular release period later on. Table 1 shows a small portion of this information. Preior to calculate the developers' areas in secion 4.4 we require this information.

@@

A relatively straightforward discipline of Linux Kernel Development team is followed with regard to the merging of patches for each release \cite{linux_kernel}.  At the beginning of each development cycle, the "merge window" is said to be opened.  At that time, code which is deemed to be sufficiently stable (and which is accepted by the development community) is merged into the mainline kernel. The bulk of changes for a new development cycle (and all of the major changes) will be merged during this time, at a rate approaching 1,000 changes ("patches," or "changesets") per day. Strategically linux starts a new kernel release with the merging to get their development branch ready to start development for the next release. So there are two main segments in a release period, one is merge window or merge period (MP) where all the developments works get merged together from different branches. So we can say this is the main development period of a release. A merge period remains opened exactly for two weeks \cite{linux_kernel}, many developers are working in many other branches to develop new features once they are ready to push it to the main kernel then they wait for the upcoming merge window and once it's merge period they all contribute their works for the next release during this Merge Period. Some developers who are mainly responsible to merge new developments verify and accept the new changes and merge them into the main kernel. Another perod in a release cycle is the period of candidate releases which gets opened once the merge window is closed where mainly the bug fixing of minor enhancement works are to be worked on and pushes by small candidate releases. Once all the fixing and enhancements are done for all the codes that had been merged into the main kernel, the release is published. We are calling this period of development as the release period (RP) because no new feature development works do not get accepted during this time, it is just the period prior to publishing a new release.

\subsubsection{Developers' Area in a Merge Period}
Merge Perid can be determined from the previous release date (of any of major, minor or micro releases) to the date of rc-1 (i.e.\ date of the first candidate release). For example if the date of release for ``linuxv2.6.12''  is 2005-06-17 and date of pushing the rc-1 for the next release ``linuxv2.6.13'' is 2005-06-29 then this time period of of 12 days is going to be being called the merge window. Thousands of commits get pushed in the main branch during this period. Table 2 shows merge periods of some releases.

\begin{table}[ht]
\caption{Release Merging Periods}  % title of Table
\centering 						% used for centering table
\begin{tabular}{c c c c}				% centered columns (4 columns)
\hline\hline						%inserts double horizontal lines
% table heading
Release 			& Start Date		& End Date \\ [0.5ex]
\hline 							% single horizontal line
% inserting body of the table
linuxv2.6.13		& 2005-06-17	& 2005-06-29 \\
linuxv2.6.14		& 2005-08-28	& 2005-09-12 \\
linuxv2.6.15		& 2005-10-27	& 2005-11-11 \\
linuxv2.6.16		& 2006-01-02	& 2006-01-17 \\
linuxv2.6.17		& 2006-03-20	& 2006-04-02 \\
linuxv2.6.18		& 2006-06-17	& 2006-07-06 \\
linuxv2.6.19		& 2006-09-19	& 2006-10-04 \\
linuxv2.6.20  		& 2006-11-29	& 2006-12-13 \\
linuxv2.6.21		& 2007-02-04	& 2007-02-20 \\
linuxv2.6.22		& 2007-04-25	& 2007-05-12 \\
linuxv2.6.23		& 2007-07-08	& 2007-07-22 \\
linuxv2.6.24		& 2007-10-09	& 2007-10-23 \\
linuxv2.6.25		& 2008-01-24	& 2008-02-10 \\
[1ex]							% adds vertical space
\hline 							% inserts single line
\end{tabular}
\label{table:nonlin} 				% is used to refer this table in the text
\end{table}

@@

\begin{table}[ht]
\caption{Developers' Areas in Merge Period}  % title of Table
\centering 						% used for centering table
\begin{tabular}{c c c c}				% centered columns (4 columns)
\hline\hline						%inserts double horizontal lines
% table heading
Author		& File Path			& Commits		& Churn \\ [0.5ex]
\hline 							% single horizontal line
% inserting body of the table
D. S. Miller	& include/.../pci.h		& 1				& 8 \\
V. Hanquez	& arch/.../cpu.c		& 1				& 14 \\
A. Bunk		& drivers/.../shmem.c	& 1				& 2 \\
J. Juhl		& arch/.../generic.c		& 1				& 3 \\
[1ex]							% adds vertical space
\hline 							% inserts single line
\end{tabular}
\label{table:nonlin} 				% is used to refer this table in the text
\end{table}

In this table we see in a merge period within a release cycle D. S. Miller has made change in pci.h file only once where he has made 8 churns. In the similar approach we calculate developers' areas in release period and also for merge period.

\subsubsection{Developers' Areas in Release Period}
As like merge period we can extract developers working in code files during release period. Release Period here in Linux Kernel development process is being considered as since the date of rc1 release (i.e.\ first release candidate opening the gate for the development of the next release) to the date of the next stable release. For example if the date of rc1 for ``linuxv2.6.13'' after releasing ``linuxv2.6.12'' is 2005-06-29 and the date of publishing the release ``linuxv2.6.13'' is 2005-08-28 then we are calling this 60 days of development period for the release as the Release Period. Table 4 is showing some development periods of different releases.

@@

\subsection{Finding Code Ownership}
So many developers are working from different development communities throught different development branches. We tried to investigate the "code ownership" \cite{mockus_case_study} evolved in Linux Kernel development. Here code ownership is going to be represented as a percentage of making churns by a developer with respect to the total churns made to a particular file. From the data that we have for complete releases (``linuxv2.6.13'' to ``linuxv3.11'') we found 70210 distinct files that have been churned by developers. We removed those files where no developer made any churn at all during these 39 releases. We investigate the total number of churns for each and every file, we also investigate how many developers contributed for these commits and changes. This information yet does not tell about the ownership, we need to find out ownership for each developer in different development areas in a release cycle. We already have the collection of developers working in merge period and release period within a release cycle and stored them with corresponding files they have worked on, how many commits a developer made to a particular file and how many changes he/she has made. We now update those information with the ownership information. Equation 1 is the base for calculating ownership.
\begin{equation}\omega=\frac{n}{N}*100\end{equation}
Where $\omega$ is the ownership and n is the number of total changes made by an atuthor of a file, N is the total number of changes made for a particular file.
To determine a developer be a owner of a file we are considering that if the developer makes changes more than 80\% of the total churn that has been made by all contributors so far for that file, then that developer can be called as a owner of that file, similarly as C. Gutwin did \cite{gutwin_awareness} for finding out the main developers' group for a project. We calculate code ownership for both Merge Period and Release Period. We also canculate general ownership for all the developers for all the files they have worked ever. Table 6 represents a part of the data which shows the ownerships of the developers in each release cycle.

>>>>

\title{On Reallocation of Resources during Releases \titlenote{Pelmission to make digital or hard copies of all or part of this work for
personal or classrooms use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page. To copy, to republish, to post on servers or to redistribute to lists,
requires prior specific permission and/or a fee. Copyright \copyright 2013}}
\numberofauthors{2}
\author{
% 1st. author
\alignauthor
Md Tajmilur Rahman\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \affaddr{438-932-2288, +1}\\
       \email{mdt\_rahm@encs.concordia.ca}
% 2nd. author
\alignauthor
Peter C. Rigby\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \affaddr{514-848-2424, +1}\\
       \email{peter.rigby@concordia.ca}
}
\date{20 Oct 2013}
\maketitle
\begin{abstract}
Software projects often suffer from lesions in our software industries due to disruptive events what may occur naturally or because of managerial directions. Keeping this in mind we have taken an approach to examine the development process of a large software project to understand the nature and behavior of the process and distribution of resources within the different areas of the system. By using the historical data from the distributed version control (DVC) system used by that large project we quantify aspects of developers' distribution, code ownership, different release periods, focus on the areas of the code-base, percentage of working with owned files during different periods withing a release cycle for this project. This study reveals a unique process performing well extraction of data from Git for the releases of such a large project by which we can perceive significant points of information that can lead us to further decisional analysis aiming novel solutions to the disruptive events for the practitioners of software development.

\end{abstract}
\category{K.6.3}{Software Management}{Software Development, Software Resource Management, Resource Reallocation}
\terms{Experiment, Human Factors, Resource Management, Reallocation}
\keywords{Resource Reallocation, Code Ownership, Defect Density, Software Releases} % NOT required for Proceedings
\section{Introduction}
Software projects are notorious for going over budget and schedule. Rush periods are often get seen before a major release that turn the developers into dinosaurs as Frederick Brooks likens in his benchmark study "The Mythical Man Month" \cite{brooks_mythical}. This "Rush To Release (RTR)" can be prompted either by external forces such as decisions by management to include new features in the release or to release earlier to beat a competitor. Alternatively, the rush may simply be due to inappropriate or unrealistic scheduling. Whatever the reason is it is an obvious. Regardless of the causes, the rush to release stresses developers and often requires developers to work on unusual, high priority or critical areas of the system. In this paper we study how RTR may effect project organization and introduce technical debt. We want to propose a method to observe, analyze and summarize the results found around different parts of a release cycle. We intend to infer behavior of the developers related to the process around release time. The key research questions that we expect to answer with our methodology are described in section 1.1.

During the merge window, developers are free to provide any change that is ready for release. During the release candidate period, we would expect to see fixes on the files that were changed during the previous merge window. We measure this by comparing the Jaccard distance of merge window with the subsequent rc period, compared with the merge window and subsequent merge window.

@@

We have organized this paper as follows. In Section 2, we describe some background and motivations followed by the summaries of related works in Section 3. Section 4 will describe methodology of our work and explanation of the data we are working with for this research. In section 5 we will discuss the results of the findings. We are describing our limitations and threats to validity in section 6 and finally section 7 will conclude the paper.

@@

\renewcommand{\labelenumi}{RQ\theenumi:}
\begin{enumerate}
\item What is the release process used by the project? \newline
We will be giving a qualitative description of Linux Kernel development in brief to give anser to this question.
\item How many developers are allocated throughout different segments of a release cycle? What significant difference can be observed between the segments of a release cycle? \newline
Answer to this question will give us a big picture of Linux Kernel development community as well as the contribution of developers in different parts within a stable release.
\item Do developers work on different areas of the system around the time of release? \newline
We want to see if developers are looking at different types of files in the code-base around the time of release which is being considered as the release period or rc period. We are going to introduce an algorithm to traverse through the git DAG to find out commits associated to a particular release. This can be useful for any other data collected from distributed version control system like git. We are expecting this would be the RTR in a release cycle which may introduce technical debts.
\item Do developers work in others' code in a high proportion during the rush period?\newline
Developers work on various files during release period. We will try to find out what is the percentage of dealing with files that developers own or they are most familiar to. This will give us a sense of the nature of work during release time.
\item How long a commit takes to get pushed into the main kernel? Does it differ in different periods of release?\newline
In distributed development environment developers work in different branches. They commit their codes to the branches they are associated to. While core development is going on developers do not directly push their changes to the main branch. We would like to see how much time it takes for a commit to the local branch to get pushed to the main branch. We also should understand how long it takes for regular development commits and how long it takes for the commits made around the release period.
\item Are there certain areas of the system that receive increased attention (i.e.\ do developers focus on a smaller set of files around releases)? \newline
To answer this question we will look into the files getting churned extremely high in number. We will try to figure out the difference between those highly churned or highly focused sets of files in different releases to understand if they require constantly high focus or they are only for one particular release.
\end{enumerate}

@@

\section{Related Works}
Hindle worked on release pattern discovery via partitioning \cite{hindle_release_pattern} to propose an approach to characterize a project's behavior around the time of major and minor releases while we are trying to study the behavior of the development resources around the time of release. In this research they proposed a method of observing, analyzing and summarizing the results of metrics of revisions found near releases. They have characterized a project's behavior around the time of major and minor releases. This is done by partitioning the observed activities like the artifact check-ins around the dates of major and minor releases, then look for reasonable patterns. Hindle divided the revisions in each release in 4 different classes, Source Code, Testing, Building, and Documentations. On the other hand Cook did an interesting job, he inserted sensors and monitors into the development process but Hindle analyzed the data to understand what happened in the past \cite{cook_automating}. Basically Hindle worked in a reverse way than Cook did. Another research work we would like to mention was done by  Damian where they have worked on the role of domain knowledge and cross functional communication among the OOS development teams \cite{damian_domain}. Posnett did some dual ecological measures of focus in software development \cite{posnett_ecological}. Posnett's measure was for the more general view that unifies developer focus and artifact ownership. He analyzed i) developer artifact contribution to network to a predator-prey food web ii) drew upon ideas from ehology to produce a novel and iii) conceptually unified view of measuring focus and ownership. Another study was done by F. Rahman about the authorship of the code-bases in OSS development \cite{rahman_ownership}.

\section{Methodology and Data}
This section presents our methodology for discovering information which can give us the idea to get the answers to our research questions. We have collected the development history data of Linux kernel. It is a data source containing the the historical data of Linux kernel development since 2005. We are going to present the steps involved in this process and then we will follow up with an application of our methodology in a case study. In order to address our research questions, we obtain key measures of project evolution from the archival data collected from the distributed version control (DVC) system of Git where all information of the OSS project ``Linux Kernel Development'' is recorded in electronic form. Many other OSS projects archive similar data, so the techniques used here can be replicated on any such project. We used the data elements extracted from the archival source to construct a number of measures on the commit log records to understand the behavior of the developers.

@@

\subsubsection{Explanation}
A Linux kernel development release version having maximum length of information does look like \textbf{linuxvA.B.C-rcP} where A, B, C, P are numeric. If a release version looks like ``linuxv2.6.13'' it tells us that this particular release is the 13th minor release in the series of the 6th major release of kernel version 2. ``linuxv2.6.13\textbf{-rc1}'' says that after the 12th stable minor release had published, development for the next release has been started and rc1 is the first candidate on the way to the next release. It also ensures that next release is going to be another minor release which is ``linuxv2.6.13''. We have last couple of release candidates for ``linuxv2.6.12'' and first couple of release candidates for ``linuxv3.12'' which is not useful for our calculation. So we are having 340 releases including major, minor, micro and rc; 39 stable releases from ``linuxv2.6.11'' to ``linuxv3.11'' having complete data with us. We have total 400441 distinct commits and for 372943 of them have change information for 77082 different files in our hand where we have 14599 developers' working history among 14621 distinct developers. Surprisingly 14569 individual files didn't have any change (addition or deletion) but was being committed at least once. So we found them in our released wise datasets.

@@

\subsection{Finding Commits for Releases}
We have to sort out and accumulate all the commits together those are belonging to any of the releases. We have all 400441 commits in our hand but though we can't plot them into a single time domain to assign different set of commits to different release periods because development goes in parallel on different branches in git, so commits also go in parallel.
\begin{figure}
\begin{center}
\includegraphics[height=2.5in,width=2.5in]{gitdag.png}
\caption{\small \sl Branching in git DAG for linux kernel development.}
\end{center}
\end{figure}
If there are multiple branches, surely there are multiple merges and the commits prior to a crisscross merge are all ancestors of the heads of all subsequent branches \cite{bird_git}. Figure 1 shows how a git DAG looks like having so many crisscross merges in a busy branching snap-shot of the git DAG of Linux Kernel development.

So it is possible to make a commit months ago from today while several releases may have already been made in the mean time but that commit can belong to any future release. Commits that are being made to a branch may get merged with other branch or the head branch during the release period. This is why we had to track all the way down the trees for each of the 340 releases in the git DAG to find out which commit contributes to which release. We could do this using Git DAG Traversing (GDT) algorithm as shown in Algorithm 1. After applying GDT we collected 381152 out of 400441 commits those are well organized in the git DAG of Linux Kernel development. As we mentioned earlier that we have last couple of release candidates for ``linuxv2.6.12'' and first couple of release candidates for ``linuxv3.12''. We are chopping them out because we don't have complete commit records for version 2.6.12-rc releases as well as 3.12-rc releases. Out of 400441 commits 19289 are belonging to those releases. Therefore, finally we have 381152 commits for 331 releases in our database to progress our work on.
\begin{algorithm}
\caption{GDT: to find Commits within Releases}
\begin{algorithmic}[1]
\REQUIRE
\STATE commit id for all releases
\ENSURE
\FOR{all commits}
\STATE start from new commit, release R
\STATE store commit in ``git\_commit\_release''
\IF{has parent commit}
	\FOR{each parent commit}
		\IF{if parent commit has a release tag}
			\STATE repeat step 1
		\ELSIF{parent commit already stored}
			\STATE compare release versions
			\IF{release version not smaller than R}
				\STATE update existing\_release $\gets$ R
			\ENDIF
		\ELSE
			\STATE store parent commit in ``git\_commit\_release''
			\STATE commit id $\gets$ parent commit id
			\STATE repeat step 3
		\ENDIF
	\ENDFOR
\ELSIF{no parent commit}
	\STATE repeat step 1
\ENDIF
\ENDFOR
\end{algorithmic}
\end{algorithm}
\subsection{Time-Span of Releases}
Another information that we require is what are the durations of the releases of Linux kernel development. To achieve that we joined the table ``git\_refs\_tags'' where all the releases are stored including the exploited release numbers and the table ``git\_commit'' to give us all the dates of commits for each and every releases, so that we can easily find out the duration between two consecutive releases (including release candidates ``rc''). This information will significantly help us because we need to see what is the development period and what are the release periods within a release cycle. We also need to understand how developers are working, what is the impact of their changes made in the code-base during a particular release period later on. Table 1 shows a small portion of this information. Prior to calculate the developers' areas in section 4.4 we require this information.

@@

A relatively straightforward discipline of Linux Kernel Development team is followed with regard to the merging of patches for each release \cite{linux_kernel}. At the beginning of each development cycle, the "merge window" is said to be opened. At that time, code which is deemed to be sufficiently stable (and which is accepted by the development community) is merged into the mainline kernel. The bulk of changes for a new development cycle (and all of the major changes) will be merged during this time, at a rate approaching 1,000 changes ("patches," or "change-sets") per day. Strategically Linux starts a new kernel release with the merging to get their development branch ready to start development for the next release. So there are two main segments in a release period, one is merge window or development period (DP) where all the developments works get merged together from different branches. So we can say this is the main development period of a release. A development period remains opened exactly for two weeks \cite{linux_kernel}, many developers are working in many other branches to develop new features once they are ready to push it to the main kernel then they wait for the upcoming merge window and once it's development period they all contribute their works for the next release during this Development Period. Some developers who are mainly responsible to merge new developments verify and accept the new changes and merge them into the main kernel. Another perod in a release cycle is the period of candidate releases which gets opened once the merge window is closed where mainly the bug fixing of minor enhancement works are to be worked on and pushes by small candidate releases. Once all the fixing and enhancements are done for all the codes that had been merged into the main kernel, the release is published. We are calling this part of a release cycle as the release period (RP) because no new feature development works get accepted during this time, it is just the period prior to publishing a new release.

\subsubsection{Developers' Area in a Development Period}
Development Period can be determined from the previous release date (of any of major, minor or micro releases) to the date of rc1 (i.e.\ date of the first candidate release). For example if the date of release for ``linuxv2.6.12'' is 2005-06-17 and date of pushing the rc1 for the next release ``linuxv2.6.13'' is 2005-06-29 then this time period of of 12 days is going to be called as the merge window or development period. Thousands of commits get pushed in the main branch during this period. Table 2 shows the development periods of some releases.

\begin{table}[ht]
\caption{Development Periods (Merge Window)}  % title of Table
\centering 						% used for centering table
\begin{tabular}{c c c c}				% centered columns (4 columns)
\hline\hline						%inserts double horizontal lines
% table heading
Release 			& Start Date		& End Date \\ [0.5ex]
\hline 							% single horizontal line
% inserting body of the table
linuxv2.6.13		& 2005-06-17	& 2005-06-29 \\
linuxv2.6.14		& 2005-08-28	& 2005-09-12 \\
linuxv2.6.15		& 2005-10-27	& 2005-11-11 \\
linuxv2.6.16		& 2006-01-02	& 2006-01-17 \\
linuxv2.6.17		& 2006-03-20	& 2006-04-02 \\
linuxv2.6.18		& 2006-06-17	& 2006-07-06 \\
linuxv2.6.19		& 2006-09-19	& 2006-10-04 \\
linuxv2.6.20  		& 2006-11-29	& 2006-12-13 \\
linuxv2.6.21		& 2007-02-04	& 2007-02-20 \\
linuxv2.6.22		& 2007-04-25	& 2007-05-12 \\
linuxv2.6.23		& 2007-07-08	& 2007-07-22 \\
linuxv2.6.24		& 2007-10-09	& 2007-10-23 \\
linuxv2.6.25		& 2008-01-24	& 2008-02-10 \\
[1ex]							% adds vertical space
\hline 							% inserts single line
\end{tabular}
\label{table:nonlin} 				% is used to refer this table in the text
\end{table}

@@

\begin{table}[ht]
\caption{Developers' Areas in Development Period}  % title of Table
\centering 						% used for centering table
\begin{tabular}{c c c c}				% centered columns (4 columns)
\hline\hline						%inserts double horizontal lines
% table heading
Author		& File Path			& Commits		& Churn \\ [0.5ex]
\hline 							% single horizontal line
% inserting body of the table
D. S. Miller	& include/.../pci.h		& 1				& 8 \\
V. Hanquez	& arch/.../cpu.c		& 1				& 14 \\
A. Bunk		& drivers/.../shmem.c	& 1				& 2 \\
J. Juhl		& arch/.../generic.c		& 1				& 3 \\
[1ex]							% adds vertical space
\hline 							% inserts single line
\end{tabular}
\label{table:nonlin} 				% is used to refer this table in the text
\end{table}

In this table we see in a development period within a release cycle D. S. Miller has made change in pci.h file only once where he has made 8 churns. In the similar approach we calculate developers' areas in release period and also for development period.

\subsubsection{Developers' Areas in Release Period}
As like development period we can extract developers working in code files during release period. Release Period here in Linux Kernel development process is being considered as since the date of rc1 release (i.e.\ first release candidate opening the gate for the development of the next release) to the date of the next stable release. For example if the date of rc1 for ``linuxv2.6.13'' after releasing ``linuxv2.6.12'' is 2005-06-29 and the date of publishing the release ``linuxv2.6.13'' is 2005-08-28 then we are calling this 60 days of development period for the release as the Release Period. Table 4 is showing some development periods of different releases.

@@

\subsection{Finding Code Ownership}
So many developers are working from different development communities throughout different development branches. We tried to investigate the "code ownership" \cite{mockus_case_study} evolved in Linux Kernel development. Here code ownership is going to be represented as a percentage of making churns by a developer with respect to the total churns made to a particular file. From the data that we have for complete releases (``linuxv2.6.13'' to ``linuxv3.11'') we found 70210 distinct files that have been churned by developers. We removed those files where no developer made any churn at all during these 39 releases. We investigate the total number of churns for each and every file, we also investigate how many developers contributed for these commits and changes. This information yet does not tell about the ownership, we need to find out ownership for each developer in different development areas in a release cycle. We already have the collection of developers working in development period and release period within a release cycle and stored them with corresponding files they have worked on, how many commits a developer made to a particular file and how many changes he/she has made. We now update those information with the ownership information. Equation 1 is the base for calculating ownership.
\begin{equation}\omega=\frac{n}{N}*100\end{equation}
Where $\omega$ is the ownership and n is the number of total changes made by an atuthor of a file, N is the total number of changes made for a particular file.
To determine a developer be a owner of a file we are considering that if the developer makes changes more than 80\% of the total churn that has been made by all contributors so far for that file, then that developer can be called as a owner of that file, similarly as C. Gutwin did \cite{gutwin_awareness} for finding out the main developers' group for a project. We calculate code ownership for both Development Period and Release Period. We also calculate general ownership for all the developers for all the files they have worked ever. Table 6 represents a part of the data which shows the ownerships of the developers in each release cycle.

\begin{table}[ht]
\caption{Code Ownership in Release Period}  % title of Table
\centering 						% used for centering table
\begin{tabular}{c c c c}				% centered columns (4 columns)
\hline\hline						%inserts double horizontal lines
% table heading
Author 				& Linuxv		& Path				& $\omega$ (\%) \\ [0.5ex]
\hline 							% single horizontal line
% inserting body of the table
Thomas Gleixner		& 2.6.24		& .../numa\_64.h	& 0.0000\\
Jeff Kirsher			& 3.1		& .../ethtool.c		& 0.0000\\
Sathya Perla			& 2.6.29		& .../hwlib.h		& 100.00\\
Roel Kluin			& 2.6.24		& .../innovator.h 	& 100.00\\
David S. Miller		& 2.6.24		& .../visasm.h 		& 100.00\\
Stephen Hemminger	& 2.6.24		& .../qla3xxx.c	 	& 12.083\\
Randy Dunlap			& 2.6.24		& ...\_core.c 		& 100.00\\
[1ex]							% adds vertical space
\hline 							% inserts single line
\end{tabular}
\label{table:nonlin} 				% is used to refer this table in the text
\end{table}
The example given above is calculated for finding ownerships of developers worked for different files in different releases. Ownership is calculated for all the developers in release periods as well as in development period. In addition we have calculated which file is owned by which developer and also the percentage of working with owned files during development and release period separately. These extractions will help us drawing the results from our findings to answer our 3rd and 4th research questions.
